Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-03-04 doc/tfp-2011/main.tex
related work

\paragraph{Related Work.}

There are several implementations of Google's Map/Reduce in functional
programming languages. But \of{Functory} was just inspired by Google's
Map/Reduce and is not exactly a Map/Reduce implementation. The
simplest difference comes from the very fact that \of{Functory} does not
operate on key/value pairs. In order to compare and better distinguish
\of{Functory} from other work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item \textit{Distributed Functional Languages (DFLs)} - functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
  Haskell, Erlang, etc.
\item \textit{Libraries for existing functional languages} - that
  could be readily used in order to avoid implementing details like
  task distribution, fault-tolerance, socket programming, etc.
\end{enumerate}

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for transmitting computations. However, it does not
provide ready-made language features for fault tolerance, which is
indispensable in a distributed setting. The user has to include code
for fault tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
programming language for distributed computing, specialized for web
programming. It provides primitives for transferring control between
the client and the server, as well as low-level primitives for
marshalling the data. In particular, it achieves type-safe
marshalling. As in the case before, ML5 is a programming language that
offers primitives for code mobility, and the code for distribution of
computation and fault-tolerance has to be included by the user. It is
worth pointing out that an existing type-safe marshalling library
could still be used with \of{Functory}. Basically, the task will be
transmitted along with the corresponding type and on unmarshalling, it
could be matched for the correct type.
\\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
Erlang~\cite{erlang} is a programming language which has features for
distribution and fault tolerance. In particular, it has features for
task distribution and is more well-known for its rich error detection
primitives and the ability to support hot-swapping. The error
detection primitives of Erlang allow nodes to monitor processes in
other nodes and also facilitate automatic migration of tasks in failed
nodes to recovered or active nodes. CamlP3l~\cite{camlP3l} mixes the
features of functional programming with predefined patterns for
parallel computation to offer a parallel programming
environment. Again, it is a programming language offering primitives
for distributing computation to parallel processes and also to merge
the results from parallel executions. The above mentioned languages
are programming languages themselves, our library, \of{Functory},
could be implemented using any of the above. The main goal of
\of{Functory} is to provide the users of an existing functional
programming language high-level user-friendly APIs that hides the
messy details of task distribution and fault-tolerance, enabling a
ready-to-use library for efficient utilization of computing resources.

We now turn to distributed computing libraries for general purpose
functional languages and weed out the distinguishing features of
\of{Functory} from other similar libraries.

PlasmaMR~\cite{plasma} is an implementation of Google's MapReduce on a
distributed file system PlasmaFS. It is able to use PlasmaFS to its
advantage - the ability of the file system to handle large files and
query functions of the functions that implement data locality to
optimize network traffic. However, Plasma MapReduce does not support
fault tolerance which is indespensible in any distributed computing
application. The Disco project~\cite{disco} is an implementation of
Google's MapReduce in Erlang~\cite{erlang}. Another MapReduce
implementation in \Ocaml\ is Yohann Padioleau's MapReduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing. Our library is not an impementation of
Google's MapReduce. Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a more generic API together with fault tolerance.

@@

  The derived API presented in Section~\ref{sec:derived} is adapted to
  deal with the absence of closures. Exactly as the \of{compute}
  function, each API now takes two forms, one for the master and
  another for the workers. For example, \of{map_fold_ac} takes the
  following forms.
  \begin{ocaml}
    Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit
    -> unit Master.map_fold_ac : 'b -> 'a list -> 'b
  \end{ocaml}
  It is the responsibility of the user to ensure type consistency
  between master and workers.

  \paragraph{\of{Mono}.} When master and workers are compiled using
  different versions of \Ocaml, we can no longer use marshalling.  As
  in the previous case, we split \of{compute} into two functions, one
  for master and one for workers. In addition, values transmitted over
  the network can only be strings. The signature thus takes the
  following form.
  \begin{ocaml}
    Worker.compute : (string -> string) -> unit -> unit Master.compute
    : master:(string * 'c -> string -> (string * 'c) list) -> (string
    * 'c) list -> unit
  \end{ocaml}
  Any other datatype for tasks should be encoded to/from strings. This
  conversion is left to the user.  Note that the second component of
  each task is still polymorphic (of type \of{'c} here), since it is
  local to the master.

>>>>

\paragraph{Related Work.}
In order to compare and better distinguish
\of{Functory} from other work with related goals and motivations, we can
broadly classify the related work in this domain into:
\begin{enumerate}
\item \textit{Distributed Functional Languages (DFLs)} --- functional
  languages that provide built-in primitives for
  distribution. Examples include ML5, \JoCaml\, Glasgow Distributed
  Haskell, Erlang, etc.
\item \textit{Libraries for existing functional languages} --- that
  could be readily used in order to avoid implementing details like
  task distribution, fault-tolerance, socket programming, etc.
\end{enumerate}

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for transmitting computations. However, it does not
provide ready-made language features for fault tolerance, which is
indispensable in a distributed setting. The user has to include code
for fault tolerance, as already demonstrated in some \JoCaml\
library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
programming language for distributed computing, specialized for web
programming. It provides primitives for transferring control between
the client and the server, as well as low-level primitives for
marshalling the data. As in the case before, ML5 is a programming language that
offers primitives for code mobility, and the code for distribution of
computation and fault-tolerance has to be included by the user. 
On the contrary, ML5 implements type-safe marshalling and \functory\
does not, though an existing type-safe marshalling library
could be used with \functory. Basically, the task will be
transmitted along with the corresponding type and on unmarshalling, it
could be matched for the correct type.
\\ \textbf{[Glasgow Distributed Haskell: TODO]}\\
Erlang~\cite{erlang} is a programming language which has features for
distribution and fault tolerance. In particular, it has features for
task distribution and is more well-known for its rich error detection
primitives and the ability to support hot-swapping. The error
detection primitives of Erlang allow nodes to monitor processes in
other nodes and also facilitate automatic migration of tasks in failed
nodes to recovered or active nodes. CamlP3l~\cite{camlP3l} mixes the
features of functional programming with predefined patterns for
parallel computation to offer a parallel programming
environment. Again, it is a programming language offering primitives
for distributing computation to parallel processes and also to merge
the results from parallel executions.

All the above mentioned languages could be used to implement our
library.  The main goal of \of{Functory}, though, is to provide the
users of an \emph{existing} functional programming language high-level
user-friendly APIs that hide the messy details of task distribution
and fault-tolerance, enabling a ready-to-use library for efficient
utilization of computing resources.
We now turn to distributed computing libraries for general purpose
functional languages and weed out the distinguishing features of
\of{Functory} from other similar libraries.

There are several implementations of Google's Map/Reduce in functional
programming languages. But \of{Functory} was just inspired by Google's
Map/Reduce and is not exactly a Map/Reduce implementation. The
simplest difference comes from the very fact that \of{Functory} does not
operate on key/value pairs. 
% TODO other difference = absence of data locality
PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's Map/Reduce on a
distributed file system PlasmaFS. It is able to use PlasmaFS to its
advantage --- the ability of the file system to handle large files and
query functions of the functions that implement data locality to
optimize network traffic. However, PlasmaMR does not support
fault tolerance which is indespensible in any distributed computing
application. 
% The Disco project~\cite{disco} is an implementation of
% Google's MapReduce in Erlang~\cite{erlang}.
Another Map/Reduce
implementation in \Ocaml\ is Yohann
Padioleau's~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing. Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a more generic API together with fault tolerance.

@@

  The derived API presented in Section~\ref{sec:derived} is adapted to
  deal with the absence of closures. Exactly as the \of{compute}
  function, each API now takes two forms, one for the master and
  another for the workers. For example, \of{map_fold_ac} takes the
  following forms.
  \begin{ocaml}
  Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
  Master.map_fold_ac : 'b -> 'a list -> 'b
  \end{ocaml}
  It is the responsibility of the user to ensure type consistency
  between master and workers.

  \paragraph{\of{Mono}.} When master and workers are compiled using
  different versions of \Ocaml, we can no longer use marshalling.  As
  in the previous case, we split \of{compute} into two functions, one
  for master and one for workers. In addition, values transmitted over
  the network can only be strings. The signature thus takes the
  following form.
  \begin{ocaml}
  Worker.compute : (string -> string) -> unit 
  Master.compute : 
    master:(string * 'c -> string -> (string * 'c) list) -> 
    (string * 'c) list -> unit
  \end{ocaml}
  Any other datatype for tasks should be encoded to/from strings. This
  conversion is left to the user.  Note that the second component of
  each task is still polymorphic (of type \of{'c} here), since it is
  local to the master.
